---
title: "Analysis of Animal Rescues (tidytuesday)"
author: "Tóth Mercédesz"
output: html_document
editor_options: 
  chunk_output_type: console
---
# 🐾 About the dataset 🐾

The dataset includes data about the London Fire Brigade's **animal rescues** since **January 2009**. As provided in the official dataset description, in **2020**, there was a **20% increase** in rescues compared to 2019. The biggest increase was observed among **non-domestic animals**.

Codebook available [here.](https://github.com/rfordatascience/tidytuesday/blob/main/data/2021/2021-06-29/readme.md)

# 🐾 Goals of this analysis 🐾

In this analysis, I will:
- prepare the dataset, compute variables and visualize them;  
- confirm the rise in rescues by plotting;  
- compare the number of domestic and non-domestic animal rescues in 2020 by plotting and t-test;  
- look for correlations between variables of interest;  
- investigate whether non-domestic animals' rescue costs more money than domestic animals' rescue by fitting a linear regression model;  
- investigate what other factors influence rescue cost with a more complex linear regression model;  
- compare the simple and the more complex models' performance.

In both models, the **outcome** variable will be incident_notional_cost.

In the more complex model, I will include the following **predictors**:  
- day_or_night: computed from date_time_of_call (0: day, 1: night);  
- pump_count: number of trucks needed for the rescue;    
- pump_hours_total: length of rescue operation;  
- domestic_wild: computed from animal_group_parent (0: domestic; 1: wild);  
- borough_inner_outer: computed from borough based on [Wikipedia](https://en.wikipedia.org/wiki/London_boroughs) (0: inner; 1: outer).

See the whole repository made for this project [here.](https://github.com/tothmercedesz2002/R-course-final)

![*Photo by Scott Walsh on Unsplash*](https://images.unsplash.com/photo-1500479694472-551d1fb6258d?q=80&w=2070&auto=format&fit=crop&ixlib=rb-4.0.3&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D)

# 🐾 Setup 🐾

## Loading packages

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(skimr)
library(correlation)
library(performance)
library(broom)
library(car)
library(lmtest)
library(ggcorrplot)
library(lm.beta)
library(ggfortify)
library(lubridate)
```

## Reading the dataset

```{r}
animal_rescues_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-06-29/animal_rescues.csv")
```

# 🐾 Preparing the dataset, computing variables and visualizing them 🐾

## First look at the dataset, filtering data and converting variables

```{r}
# Checking structure and variables' type

str(animal_rescues_raw)

# Selecting variables of interest

animal_rescues_interest <- animal_rescues_raw %>%
  select(date_time_of_call, cal_year, pump_count, pump_hours_total, incident_notional_cost, animal_group_parent, borough)

# Looking for missing data

sum(is.na(animal_rescues_interest))
which(rowSums(is.na(animal_rescues_interest)) > 0)
missing_info <- animal_rescues_interest[c(4138, 4165, 4246, 4709, 5897, 6339, 6768, 7072, 7478), ]
print(missing_info)

unique(animal_rescues_interest$pump_count)
unique(animal_rescues_interest$pump_hours_total)
unique(animal_rescues_interest$incident_notional_cost)
unique(animal_rescues_interest$animal_group_parent)

# Filtering rows with missing values and "NULL" values

animal_rescues_interest <- animal_rescues_interest %>%
  filter(!is.na(borough)) %>%
  filter(pump_count != "NULL") %>%
  filter(pump_hours_total != "NULL") %>%
  filter(incident_notional_cost != "NULL")

# Converting variables

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(
    date_time_of_call = dmy_hm(date_time_of_call),
    pump_count = as.numeric(pump_count),
    pump_hours_total = as.numeric(pump_hours_total),
    incident_notional_cost = as.numeric(incident_notional_cost),
    animal_group_parent = as.factor(animal_group_parent),
    borough = as.factor(borough)
    )

str(animal_rescues_interest)
summary(animal_rescues_interest)
```

## Computing day_or_night variable and visualization

```{r}
# Computing day_or_night variable
# 0 - day: 06:00 - 18:00
# 1 - night: 18:01 - 05:59

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(
    day_or_night = case_when(
      hour(date_time_of_call) >= 6 & hour(date_time_of_call) < 18 ~ 0,
      TRUE ~ 1)
  )

# Creating an extra variable for plotting

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(
    day_or_night_forplots = factor(
      day_or_night,
      levels = c(0, 1),
      labels = c("Day", "Night"))
  )

# Checking and comparing counts of day and night animal rescues

ggplot(animal_rescues_interest, aes(day_or_night_forplots, fill = day_or_night_forplots)) +
  geom_bar() +
  labs(
    x = "Time of day",
    y = "Number of rescues",
    title = "Animal rescues by time of day"
  ) +
  scale_fill_manual(values = c("Day" = "#cc5500", "Night" = "#000080")) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
```

## Computing domestic_wild variable and visualization

```{r}
# Computing domestic_wild variable
# 0: domestic (traditional pets and livestock)
# 1: wild (including exotic animals kept as pets)

levels(animal_rescues_interest$animal_group_parent)

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(animal_group_parent = fct_recode(animal_group_parent, 
                                          "Cat" = "cat", 
                                          "Sheep" = "Lamb", 
                                          "Bird" = "Pigeon", 
                                          "Cow" = "Bull")) 
  
levels(animal_rescues_interest$animal_group_parent)

# Domestic: "Cow", "Hamster", "Horse", "Unknown - Animal rescue from water - Farm animal", "Unknown - Heavy Livestock Animal", "Cat", "Dog", "Goat", "Sheep", "Rabbit", "Unknown - Animal rescue from below ground - Farm animal", "Unknown - Domestic Animal Or Pet"

# Wild: "Bird", "Deer", "Ferret", "Hedgehog", "Fox", "Lizard", "Snake", "Tortoise", "Budgie", "Fish", "Hedgehog", "Squirrel", "Unknown - Wild Animal" 

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(domestic_wild = ifelse(animal_group_parent %in% c(
    "Cow", "Hamster", "Horse", "Unknown - Animal rescue from water - Farm animal", 
    "Unknown - Heavy Livestock Animal", "Cat", "Dog", "Goat", "Sheep", "Rabbit", 
    "Unknown - Animal rescue from below ground - Farm animal", "Unknown - Domestic Animal Or Pet"
    ), 0, 1))

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(domestic_wild = as.factor(domestic_wild))

# Creating an extra variable for plotting

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(
    domestic_wild_forplots = factor(
      domestic_wild,
      levels = c(0, 1),
      labels = c("Domestic", "Wild"))
  )

# Checking and comparing counts of domestic and wild rescued animals

ggplot(animal_rescues_interest, aes(domestic_wild_forplots, fill = domestic_wild_forplots)) +
  geom_bar() +
  labs(
    x = "Type of animal",
    y = "Number of rescues",
    title = "Animal rescues by animal type"
  ) +
  scale_fill_manual(values = c("Domestic" = "pink", "Wild" = "purple")) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
```

## Computing borough_inner_outer and visualization

```{r}
# Computing borough_inner_outer
# 0: inner
# 1: outer

levels(animal_rescues_interest$borough)

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(borough = str_to_title(as.character(borough))) %>%
  mutate(borough = factor(borough))

levels(animal_rescues_interest$borough)

valid_boroughs <- c(
  "Barking and Dagenham", "Barnet", "Bexley", "Brent", "Bromley", "Camden", "Croydon", "Ealing",
  "Enfield", "Greenwich", "Hackney", "Hammersmith And Fulham", "Haringey", "Harrow", "Havering", 
  "Hillingdon", "Hounslow", "Islington", "Kensington And Chelsea", "Kingston upon Thames", 
  "Lambeth", "Lewisham", "Merton", "Newham", "Redbridge", "Richmond Upon Thames", "Southwark", 
  "Sutton", "Tower Hamlets", "Waltham Forest", "Wandsworth", "Westminster"
)

animal_rescues_interest <- animal_rescues_interest %>%
  filter(borough %in% valid_boroughs)
animal_rescues_interest$borough <- droplevels(animal_rescues_interest$borough)

levels(animal_rescues_interest$borough)
nlevels(animal_rescues_interest$borough)

# inner boroughs: "Camden", "Greenwich", "Hackney", "Hammersmith And Fulham", "Islington", "Kensington And Chelsea", "Lambeth", "Lewisham", "Southwark", "Tower Hamlets", "Wandsworth", "Westminster"

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(borough_inner_outer = ifelse(borough %in% c(
    "Camden", "Greenwich", "Hackney", "Hammersmith And Fulham", "Islington", "Kensington And Chelsea", "Lambeth",
    "Lewisham", "Southwark", "Tower Hamlets", "Wandsworth", "Westminster"
    ), 0, 1))

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(borough_inner_outer = as.factor(borough_inner_outer))

# Creating an extra variable for plotting

animal_rescues_interest <- animal_rescues_interest %>%
  mutate(
    borough_inner_outer_forplots = factor(
      borough_inner_outer,
      levels = c(0, 1),
      labels = c("Inner", "Outer"))
  )

# Checking and comparing counts of animal rescues in inner and outer boroughs of London

ggplot(animal_rescues_interest, aes(borough_inner_outer_forplots, fill = borough_inner_outer_forplots)) +
  geom_bar() +
  labs(
    x = "Type of borough",
    y = "Number of rescues",
    title = "Animal rescues in London boroughs"
  ) +
  scale_fill_manual(values = c("Inner" = "#50c878", "Outer" = "#db7093")) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
```

## Exploring distribution of remaining variables of interest
```{r}
# pump_count

mean(animal_rescues_interest$pump_count)
sd(animal_rescues_interest$pump_count)
min(animal_rescues_interest$pump_count)
max(animal_rescues_interest$pump_count)

ggplot(animal_rescues_interest, aes(pump_count)) +
  geom_histogram(binwidth = 1, fill = "#e6e6fa", color = "black") +
  labs(
    x = "Number of pumps",
    y = "Count",
    title = "Distribution of pump count"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Due to pump_count = 1 overwhelming the plot, here is another plot with pump_count > 1

ggplot(animal_rescues_interest %>% filter(pump_count > 1), aes(pump_count)) +
  geom_histogram(binwidth = 1, fill = "#e6e6fa", color = "black") +
  labs(
    x = "Number of pumps",
    y = "Count",
    title = "Distribution of pump count (> 1)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# pump_hours_total

mean(animal_rescues_interest$pump_hours_total)
sd(animal_rescues_interest$pump_hours_total)
min(animal_rescues_interest$pump_hours_total)
max(animal_rescues_interest$pump_hours_total)

ggplot(animal_rescues_interest, aes(pump_hours_total)) +
  geom_histogram(binwidth = 1, fill = "#b0e0e6", color = "black") +
  labs(
    x = "Hours",
    y = "Count",
    title = "Distribution of rescue length"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Due to low values overwhelming the plot, here is another plot with pump_hours_total > 3

ggplot(animal_rescues_interest %>% filter(pump_hours_total > 3), aes(pump_hours_total)) +
  geom_histogram(binwidth = 1, fill = "#b0e0e6", color = "black") +
  labs(
    x = "Hours",
    y = "Count",
    title = "Distribution of rescue length (> 3)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# incident_notional_cost

mean(animal_rescues_interest$incident_notional_cost)
sd(animal_rescues_interest$incident_notional_cost)
min(animal_rescues_interest$incident_notional_cost)
max(animal_rescues_interest$incident_notional_cost)

ggplot(animal_rescues_interest, aes(incident_notional_cost)) +
  geom_histogram(binwidth = 100, fill = "#8fbc8f", color = "black") +
  labs(
    x = "Notional cost of rescues",
    y = "Count",
    title = "Distribution of notional cost of rescues"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Due to low values overwhelming the plot, here is another plot with incident_notional_cost > 1000

ggplot(animal_rescues_interest %>% filter(incident_notional_cost > 1000), aes(incident_notional_cost)) +
  geom_histogram(binwidth = 100, fill = "#8fbc8f", color = "black") +
  labs(
    x = "Notional cost of rescues",
    y = "Count",
    title = "Distribution of notional cost of rescues (> 1000)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Since the cost of the rescue is the main focus of my analysis (I want to know what causes the extreme values), I will leave the variable as it is (no transformations).
```

# 🐾 Confirming the rise in rescues by plotting 🐾
